---
title: "Gender Based Violence Analysis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
   bookdown::word_document2:
      fig_caption: yes
      toc: true
      toc_depth: 2
      reference_docx: "../template/template.docx"
---

```{r, setup, include=FALSE, warning = FALSE, message=FALSE}
library(stringr)
library(officedown)
library(officer) 
library(knitr)
current_wd <- getwd()

if(endsWith(current_wd, "gbv-health-provider-study")){
  disc_project_wd <- current_wd
} else if (endsWith(current_wd, "/paper")) {
  disc_project_wd <- str_remove(current_wd, "/paper")   
} else {
  print("Got a WD that's not handled in the If-else ladder yet")
}

knitr::opts_chunk$set(fig.align = 'center',
                      echo = FALSE,
                      warning = FALSE, 
                      message = FALSE,
                      dpi = 300)

## SOURCE FIGURES AND TABLES
source(paste(disc_project_wd, "/code/table_1.R", sep = ""))

```

# Analysis plan

## Background
Gender-based violence (GBV) has significant negative effects on the health and wellbeing of those affected and is pervasive in Southeast Asia. Women who are victims of violence utilize healthcare more often than women who are not victims of violence. Health providers are uniquely positioned to intervene and offer support to women affected by GBV. As such, GBV interventions should consider the role of the health system. Timor-Leste has responded to the high burden of GBV in the country by enacting laws that include health providers as important actors against GBV. This systems-wide approach places emphasis on the role of the health system in responding to GBV. 

While Timor-Leste has identified the health system as essential in responding to GBV, few health providers have received training on GBV. A Lancet paper highlighted the importance of building clinicians knowledge and skill in responding to GBV, as well as strengthening their attitudes toward GBV. A 2021 meta-analysis of GBV training programs for health providers found that training may improve GBV knowledge, attitudes, identification, and self-confidence in responding to GBV 12-months post-training when compared to no training or training as usual. However, three quarters of the studies included in this analysis were U.S.-based, with very few studies from low or middle-income countries. While trainings may improve GBV-related knowledge and behaviors amongst health providers, there is evidence to suggest that single trainings are less effective in improving and maintaining knowledge than trainings with follow-up sessions for reinforcement of content. These studies highlight the effectiveness of training health providers in GBV, but a gap remains in understanding how GBV trainings for health providers can be implemented in low and middle-income countries and what impact they may produce in settings with varying cultures and context. This line of research is important to continue because it provides critical information for developing and implementing effective interventions for a health system response to GBV, particularly in low and middle-income countries and in settings where GBV is more culturally embedded.

## Aims
Aim 1:  To determine if there is significant differences in health provider knowledge, attitudes, empathy, system support, confidence, and practices in responding to GBV before and after participation in the 'Responding to GBV' curriculum in the intervention municipalities in Timor-Leste, using baseline (pre-training) and 14-month post-learning lab data.

Sub-aim 1.1: To evaluate changes in health provider knowledge, attitudes, empathy, system support, confidence, and practices in responding to GBV after participation in the  'Responding to GBV' intensive training compared to the 14-months of follow-up learning labs after the training in the intervention municipalities in Timor-Leste, using percent change from baseline to post-intensive training, and post-intensive to endline.

Aim 2:  To identify provider-level characteristics associated with change in GBV knowledge, attitudes, empathy, system support, confidence, and practice scores, including age, sex, occupation, years of professional experience, and attendance at learning labs.

## Study details

## Outcome

## Study procedures

## Statistical analysis


# Results

## Demographic information
```{r}
demographic_table %>%
  modify_caption("**Table 1. Participant Characteristics**")
```

## Scores

## Regression results