---
title: "Gender Based Violence Analysis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
   bookdown::word_document2:
      fig_caption: yes
      toc: true
      toc_depth: 2
      reference_docx: "../template/template.docx"
---

```{r, setup, include=FALSE, warning = FALSE, message=FALSE}
library(stringr)
library(officedown)
library(officer) 
library(knitr)
current_wd <- getwd()

if(endsWith(current_wd, "gbv-health-provider-study")){
  disc_project_wd <- current_wd
} else if (endsWith(current_wd, "/paper")) {
  disc_project_wd <- str_remove(current_wd, "/paper")   
} else {
  print("Got a WD that's not handled in the If-else ladder yet")
}

knitr::opts_chunk$set(fig.align = 'center',
                      echo = FALSE,
                      warning = FALSE, 
                      message = FALSE,
                      dpi = 300)

## SOURCE FIGURES AND TABLES
source(paste(disc_project_wd, "/code/table_1.R", sep = ""))
source(paste(disc_project_wd, "/code/analysis.R", sep = ""))
source(paste(disc_project_wd, "/code/plots.R", sep = ""))
source(paste(disc_project_wd, "/code/exploratory_data_analysis.R", sep = ""))
```

# Analysis plan
### Program Description
This analysis focuses on the ‘Responding to GBV’ arm of the Harmonia Activity. The goal of the ‘Responding to GBV’ curriculum was to educate health providers in GBV dynamics and shift social norms with the future intent of providing better protection for GBV victims. The training intervention took place at regional health facilities and consisted of an initial five days of intensive training, followed by monthly ‘learning lab’ sessions to be carried out at health facilities over a 14-month period. All health providers in the selected municipalities were eligible to participate in the trainings. Participants included medical doctors or nurses, as well as health facility managers, pharmacists, lab analysts, and other positions. From July to November 2021, eleven intensive trainings took place at health facilities in Ermera and Liquica, training a total of 302 healthcare workers on GBV. These five-day intensive trainings consisted of 14 two-hour modules on various GBV topics. Learning labs, which were shorter 2-hour trainings covering one GBV module each, ran for 14-months following the intensive trainings, from March 2022 to June 2023.

## Aims
Aim 1: To determine if there is significant difference in health provider knowledge, attitudes, empathy, system support, confidence, and practices in responding to GBV before and after participation in the 'Responding to GBV' curriculum in the intervention municipalities in Timor-Leste, using baseline (pre-training) and 14-month post-learning lab data.

Sub-aim 1.1: To evaluate changes in health provider knowledge, attitudes, empathy, system support, confidence, and practices in responding to GBV after participation in the  'Responding to GBV' intensive training compared to the 14-months of follow-up learning labs after the training in the intervention municipalities in Timor-Leste, using percent change from baseline to post-intensive training, and post-intensive to endline.

Aim 2: To identify provider-level characteristics associated with change in GBV knowledge, attitudes, empathy, system support, confidence, and practice scores, including age, sex, occupation, years of professional experience, and attendance at learning labs.

## Methods
### Data Collection and Questionnaire
An adaptation of the WHO curriculum evaluation tool was used to assess knowledge, attitudes, confidence, and system support. This evaluation measure consisted of 89 questions, organized into four domains related to the GBV training curriculum: knowledge, attitudes, confidence, and system support. The survey and its related domains can be viewed in the supplemental materials.

The knowledge domain included 43 questions, each assessed with responses of yes, no, or I don’t know, for a maximum possible score of 43. Examples of questions included, “there are common injury patterns associated with domestic violence,” “it is a healthcare provider’s legal duty to help the woman subjected to violence to report it to the police,” and “is depression, anxiety, or chronic stress a warning sign that a woman may have been subjected to domestic or sexual violence?”. Questions were scored as 1 for correct (“yes”) responses and 0 for incorrect (“no” or “I don’t know”) responses. Higher scores were indicative of greater GBV-related knowledge.

The attitudes domain included 29 questions measured on 4 and 5-point Likert scales for a maximum possible score of 102. Examples of questions included “domestic violence is a private matter and people outside the family should not interfere”, “it is humiliating to patients to question them about abuse”, and “it is acceptable for a man to hit his wife or girlfriend if she argues with him.” Questions were formulated in both positive and negative terms. For questions using the 5-point Likert scale, responses were scored on a scale where positively worded items (where “strongly agree” was the preferred answer) were scored as follows: strongly disagree = 0; disagree = 1; neither agree nor disagree = 2; agree = 3; strongly agree = 4. Conversely, negatively worded items (where “strongly disagree” was the desired response) had their scores reversed to account for the preferred response, with strongly disagree obtaining the highest score of 4. For questions using the 4-point Likert scale, the desirable response for all questions was “no, it is never acceptable.” In this case, responses were scored as follows: yes, it is acceptable = 0; sometimes it is acceptable or I don’t know = 1; no, it is never acceptable = 2. Higher scores were indicative of more desirable health provider attitudes toward GBV and the health provider role in responding to GBV.

The confidence domain included 10 questions measured on a 5-point Likert scale for a maximum possible score of 40. Examples of questions included asking health providers if they felt prepared to, “identify a woman who has been subjected to violence by signs and symptoms she reports”, “offer supportive statements to a women subjected to domestic violence or sexual assault”, and “to assess the immediate level of danger for a woman or child after sexual assault or domestic violence.” Responses were scored as follows: not at all prepared = 0; slightly prepared = 1; somewhat prepared = 2; sufficiently prepared = 3; and quite well prepared = 4. Higher scores were indicative of greater confidence in identifying and caring for women subjected to violence.

The system support domain included 6 questions, each assessed with responses of yes, no, or I don’t know, for a maximum possible score of 6. Examples of questions included, “I can readily look up information (e.g. either a guide or standard operating procedure) on how to manage cases of domestic violence or sexual assault” and “my supervisor supports me pro-actively asking my patients or clients about whether they are experiencing domestic violence.” Questions were scored as 1 for correct (“yes”) responses and 0 for incorrect (“no” or “I don’t know”) responses. Higher scores were indicative of a perception of greater resources and support for health providers to respond to GBV within their facility.

The Toronto Empathy Questionnaire (TEQ) was used to provide a measure of empathy for GBV survivors. The TEQ is a brief, unidimensional tool to measure empathy as a broad construct across populations.13 The tool was developed and validated in Canada, where studies demonstrated high internal consistency and reliability amongst university students.13 Subsequently, the TEQ have been administered to assess empathy in various settings, including, but not limited to, medical students in South Korea14 and Romania15, and amongst the general population in the Czech Republic.16 In these settings, the TEQ was translated into the appropriate language and adapted versions of the TEQ were shown to be more valid.14–16 Adaptations included rewording of statements and/or exclusion of survey items.13–15 To date, there is no evidence to support the TEQ’s previous use in Timor-Leste or translation into Tetum.

The empathy domain, derived from the TEQ, included 16 questions measured on a 5-point Likert scale for a total possible score of 64. Examples of questions included, “I enjoy making others feel better”, “I become irritated when someone cries”, and “it upsets me to see someone being treated disrespectfully.” Responses were scored on a scale where positively worded items (where “always” was the preferred answer) were scored as follows: never = 0; rarely = 1; sometimes = 2; often = 3; always = 4. Conversely, negatively worded items (where “never” was the desired response) had the scores reversed to account for the preferred response, with “never” obtaining the highest score of 4. Higher scores were indicative of greater self-reported empathic behavior.

Composite variables for each domain will be created by summing the scores of each survey item within their respective domain.

All participants filled out paper survey forms, and the data were subsequently entered into the RedCap data management software by HAMNASA staff. Fewer than 1% of responses were missing from the scored domains.

# Results

## Demographic information
```{r}
demographic_table <- demographic_table %>%
  modify_caption("**Table 1. Participant Characteristics**")
demographic_table
```
Matched data was obtained for 78 individuals, representing a total of 234 surveys out of the 676 conducted across the three timepoints. There were no significant demographic differences between individuals with less than three timepoints and all three timepoints, leading us to believe the smaller longitudinal sample remained representative of the intervention population. Demographic characteristics and work history of participants at the baseline are presented in Table 1. The overall sample had `r inline_text(demographic_table, variable = sex_factored, level = "Female", column= "stat_0")` female and `r inline_text(demographic_table, variable = sex_factored, level = "Male", column= "stat_0")` male participants. `r inline_text(demographic_table, variable = age_collapsed, level = "<=34", column= "stat_0")` of the sample was 34 years old or younger. The most prevalent occupational role was medical doctor, accounting for `r inline_text(demographic_table, variable = position_groups, level = "Medical doctor", column= "stat_0")`, followed by nurses at `r inline_text(demographic_table, variable = position_groups, level = "Nurse", column= "stat_0")`, and midwives at `r inline_text(demographic_table, variable = position_groups, level = "Midwife", column= "stat_0")`. On average, participants had been practicing in their clinical roles for approximately `r inline_text(demographic_table, variable = position_years_clean, column= "stat_0")` years. Only `r inline_text(demographic_table, variable = previous_training_factored, column= "stat_0")` of participants had received any previous GBV training.

## Attendance
```{r}
attendance_hist <- hist(df_wide$attendance_score_FUAT, 
                        main = "FUAT Attendance", 
                        xlab = "Number of FUAT sessions attended", 
                        ylab = "Frequency")
```

## Scores
```{r}
comparison_tbl_1 <- comparison_tbl_1 %>%
  modify_caption("**Table 2. Median participant scores at baseline and endline, by domain**")
comparison_tbl_1
```

Significant improvement in participant scores were observed from baseline (timepoint 1) to endline (timepoint 3) in each domain, as illustrated in table 2 (p <.001). From baseline to endline, median GBV knowledge improved from `r inline_text(comparison_tbl_1, variable = knowledge_overall, column= "stat_1")` to `r inline_text(comparison_tbl_1, variable = knowledge_overall, column= "stat_2")`, out of 43 possible points. The attitudes toward GBV domain improved from `r inline_text(comparison_tbl_1, variable = attitude_overall, column= "stat_1")` to `r inline_text(comparison_tbl_1, variable = attitude_overall, column= "stat_2")`, out of 102 possible points. Confidence scores that represented provider confidence in identifying and treating women subjected to GBV improved from `r inline_text(comparison_tbl_1, variable = confidence_score, column= "stat_1")` to `r inline_text(comparison_tbl_1, variable = confidence_score, column= "stat_2")`, out of 40 possible points. Empathy scores, representing provider's empathic behavior, improved from `r inline_text(comparison_tbl_1, variable = empathy_score, column= "stat_1")` to `r inline_text(comparison_tbl_1, variable = empathy_score, column= "stat_2")`, out of 64 possible points. System support scores rose from `r inline_text(comparison_tbl_1, variable = system_support_score, column= "stat_1")` to `r inline_text(comparison_tbl_1, variable = system_support_score, column= "stat_2")`, out of 6 possible points. While comparatively the practice domain had a small number of respondents (n = `r inline_text(comparison_tbl_1, variable = practice_score, column= "n")`) due to the skip logic employed in the practice questions, the results were still significant. Practice scores, measuring appropriate actions taken by providers to support women identified as victims of GBV, improved from `r inline_text(comparison_tbl_1, variable = practice_score, column= "stat_1")` to `r inline_text(comparison_tbl_1, variable = practice_score, column= "stat_2")`.

While significant improvements were seen from intervention baseline to endline, no significant differences were observed between timepoint 2 (post 5-day intensive GBV training) and timepoint 3 (endline, following 14-months of monthly learning labs).
```{r}
comparison_tbl_2 <- comparison_tbl_2 %>%
  modify_caption("**Table 3. Median participant scores at midline and endline, by domain**")
comparison_tbl_2
```

Scores plateaud between timepoints 2 and 3. Notably, the domains of system support and practices should not have improved between timepoint 1 and timepoint 2. The practices domain asks respondents to report their practices in caring for women in the past month. Those answers should be the same between timepoints 1 and 2.  Similarly, the system support domain discusses perceived support (both within facilities and outside of facilities) for providers caring for victims of GBV. While it's not impossible for these scores to have improved (e.g. "I have names and contact information of people within this facility to whom I can refer the client who discloses violence for additional counselling or psychosocial support" could have improved), not all questions should have imporved (e.g."I have a private space in the facility where I can talk to the woman confidentially about her abuse"). As such, should we throw them out of the analysis?
```{r}
scoreplot_log
```

## Regression results
```{r}

```

