---
title: "Gender Based Violence Analysis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
   bookdown::word_document2:
      fig_caption: yes
      toc: true
      toc_depth: 2
      reference_docx: "../template/template.docx"
---

```{r, setup, include=FALSE, warning = FALSE, message=FALSE}
library(stringr)
library(officedown)
library(officer) 
library(knitr)
current_wd <- getwd()

if(endsWith(current_wd, "gbv-health-provider-study")){
  disc_project_wd <- current_wd
} else if (endsWith(current_wd, "/paper")) {
  disc_project_wd <- str_remove(current_wd, "/paper")   
} else {
  print("Got a WD that's not handled in the If-else ladder yet")
}

knitr::opts_chunk$set(fig.align = 'center',
                      echo = FALSE,
                      warning = FALSE, 
                      message = FALSE,
                      dpi = 300)

## SOURCE FIGURES AND TABLES
source(paste(disc_project_wd, "/code/table_1.R", sep = ""))

```

# Analysis plan

## Background

## Aims

Aim 1: To assess change in individual health provider knowledge, attitudes, empathy, and confidence in responding to gender-based violence (GBV) before and after participation in the 'Responding to GBV' curriculum in the intervention municipalities in Timor-Leste.
  Sub-aim 1.1: To assess change in self-reported provider practices in identifying and responding to gender based violence in the past    month.

Aim 2: To assess change in advancing individual health providers’ knowledge, attitudes, and empathy related to GBV after the 14-month Learning Lab curriculum when compared to the 5-day intensive training, using baseline, post-intensive, and endline data.

## Study details

## Outcome

## Study procedures

## Statistical analysis


# Results

## Demographic information
```{r}
demographic_table %>%
  modify_caption("**Table 1. Participant Characteristics**")
```

## Scores

## Regression results