---
title: "Gender Based Violence Analysis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
   bookdown::word_document2:
      fig_caption: yes
      toc: true
      toc_depth: 2
      reference_docx: "../template/template.docx"
---

```{r, setup, include=FALSE, warning = FALSE, message=FALSE}
library(stringr)
library(officedown)
library(officer) 
library(knitr)
current_wd <- getwd()

if(endsWith(current_wd, "gbv-health-provider-study")){
  disc_project_wd <- current_wd
} else if (endsWith(current_wd, "/paper")) {
  disc_project_wd <- str_remove(current_wd, "/paper")   
} else {
  print("Got a WD that's not handled in the If-else ladder yet")
}

knitr::opts_chunk$set(fig.align = 'center',
                      echo = FALSE,
                      warning = FALSE, 
                      message = FALSE,
                      dpi = 300)

## SOURCE FIGURES AND TABLES
source(paste(disc_project_wd, "/code/table_1.R", sep = ""))

```

# Analysis plan

## Background

## Aims

Aim 1: To assess change in individual health provider knowledge, attitudes, empathy, and confidence in responding to gender-based violence (GBV) before and after participation in the 'Responding to GBV' curriculum in the intervention municipalities in Timor-Leste.
  Sub-aim 1.1: To assess change in self-reported provider practices in identifying and responding to gender based violence in the past    month.

Aim 2: To assess change in advancing individual health providers’ knowledge, attitudes, and empathy related to GBV after the 14-month Learning Lab curriculum when compared to the 5-day intensive training, using baseline, post-intensive, and endline data.

## Study details

**Design:** Pre/post intervention comparison
**Population:** Healthcare providers in Liquica and Ermera municipalities, Timor-Leste.
**Sites:** 10 community health centers (CHCs) total, 6 in Liquica and 4 in Ermera municipality. Health post staff were eligible to attend at their local CHC, with a maximum of 63 health posts in the municipalities.
**Sample size:** Timepoints 1-3 had 216, 208, and 252 respondents who consented to research respectively. Across all three timepoints, we were able to match 57 participants. 

## Outcome
Scores in six domains: 1) knowledge, 2) attitudes, 3) system support, 4) confidence, 5) empathy and 6) practices. The domains of knowledge and attitudes had four sub-domains each. Knowledge was broken up into: 1) general knowledge of GBV, 2) warning signs of GBV, 3) appropriate ways to ask about GBV and 4) helpful responses to support a woman subjected to GBV. Attitudes was broken up into: 1) general attitudes towards GBV and the health provider role, 2) acceptability for a man to his his partner, 3) attitudes towards gender roles, and 4) attitudes towards professional roles.

## Study procedures



## Data cleaning

## Statistical analysis


# Results

## Demographic information
```{r}
demographic_table %>%
  modify_caption("**Table 1. Participant Characteristics**")
```

## Scores

## Regression results