---
title: "Gender Based Violence Analysis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
   bookdown::word_document2:
      fig_caption: yes
      toc: true
      toc_depth: 2
      reference_docx: "../template/template.docx"
---

```{r, setup, include=FALSE, warning = FALSE, message=FALSE}
library(stringr)
library(officedown)
library(officer) 
library(knitr)
current_wd <- getwd()

if(endsWith(current_wd, "gbv-health-provider-study")){
  disc_project_wd <- current_wd
} else if (endsWith(current_wd, "/paper")) {
  disc_project_wd <- str_remove(current_wd, "/paper")   
} else {
  print("Got a WD that's not handled in the If-else ladder yet")
}

knitr::opts_chunk$set(fig.align = 'center',
                      echo = FALSE,
                      warning = FALSE, 
                      message = FALSE,
                      dpi = 300)

## SOURCE FIGURES AND TABLES
source(paste(disc_project_wd, "/code/table_1.R", sep = ""))
source(paste(disc_project_wd, "/code/analysis.R", sep = ""))
```

# Analysis plan

## Background

Gender-based violence (GBV) has significant negative effects on the health and wellbeing of those affected and is pervasive in Southeast Asia. GBV is defined as “harmful acts directed at an individual or a group of individuals based on their gender.”1 GBV is a broad term and can include domestic violence, sexual assault, and child abuse.2 In Timor-Leste, male-perpetrated violence against female partners is the leading form of GBV.3

In 2018, the World Health Organization (WHO) estimated that the lifetime prevalence of physical and/or sexual GBV for ever-married or partnered women 15 years and older was 26% globally, equating to approximately 641 million women.4 Globally, 10% of ever-married or partnered women 15 years and older experienced GBV within the past year.4 GBV rates within Timor-Leste vary by study, but all demonstrate higher rates of violence when compared to other countries. The WHO reported that Timor-Leste was placed in the top fourteen countries studied with regard to past year GBV, at 28% of women.4 Rates of past year and lifetime experience of GBV vary by study, but range from 28-47% and 35-59% respectively.3,5 The variability in rates may be due to how the research was conducted and reporting of GBV may be impacted by the willingness of women to disclose their experiences of GBV.2 As GBV is often under-reported, the book Gender–Based Violence & Healthcare in Timor-Leste suggests that higher estimates are likely more accurate.2

GBV has a known impact on health and wellbeing. On a global scale, GBV leads to a greater burden of disability-adjusted life years among women of reproductive age compared to well known health risk factors, such as smoking.6 Women experiencing GBV are more likely to be infected with HIV or experience depression, with 54% and 60% increased risks respectively when compared to women not experiencing GBV.6 Data from 2009/2010 Timor-Leste Demographic and Health Survey found that women who experienced combined forms of violence in Timor-Leste were three to four times more likely to have a sexually transmitted infection (STI) or STI symptoms, 76% more likely to have only two or three antenatal care visits compared to the standard of four or more, 45% more likely to have at least one of their children die, and 46% more likely to have a low birth weight baby than women who had not experienced GBV.7

Women who are victims of violence utilize healthcare more often than women who are not victims of violence.8 Health providers play a unique role in responding to GBV, with each patient interaction providing an opportunity to identify women subjected to violence and intervene. With these increased touchpoints between women who are victims of violence and the healthcare system, health providers can provide immediate and ongoing physical and mental health care to victims of GBV, referral to supportive services, as well as documentation of cases for legal purposes.8 Health providers are uniquely positioned to intervene and offer support to women affected by GBV. As such, GBV interventions should consider the role of the health system. 

Timor-Leste has responded to the high burden of GBV in the country by enacting laws that include health providers as important actors against GBV. In 2010, the government enacted the Law Against Domestic Violence, outlawing violence against family members.2 Articles 22 and 40 in the Law Against Domestic Violence describe the role of health providers specifically in responding to GBV.2 These articles state that providers are legally responsible for protecting patient confidentiality and are obligated to enter gender-based violence cases into the medical record, sharing that report with police for investigative purposes.2 The 2012 National Action Plan on Gender-Based Violence outlines a multi-sectoral approach for addressing GBV in Timor-Leste7, highlighting the role of the health sector in providing safe, confidential care to victims of GBV, with referral to additional services as necessary.9 Timor-Leste has taken a systems-wide approach to addressing its high rate of GBV, with emphasis placed on the role of the health system. 

While Timor-Leste has identified the health system as essential in responding to GBV, few health providers have received training on GBV. A Lancet paper highlighted the importance of building clinicians knowledge and skill in responding to GBV, as well as strengthening their attitudes toward GBV.8 A 2021 meta-analysis of GBV training programs for health providers found that training may improve GBV knowledge, attitudes, identification, and self-confidence in responding to GBV 12-months post-training when compared to no training or training as usual.10 However, three quarters of the studies included in this analysis were U.S.-based, with very few studies from low or middle-income countries.10 While trainings may improve GBV-related knowledge and behaviors amongst health providers, there is evidence to suggest that single trainings are less effective in improving and maintaining knowledge than trainings with follow-up sessions for reinforcement of content.11 These studies highlight the effectiveness of training health providers in GBV, but a gap remains in understanding how GBV trainings for health providers can be implemented in low and middle-income countries and what impact they may produce in settings with varying cultures and context. 

This line of research is important to continue because it provides critical information for developing and implementing effective interventions for a health system response to GBV, particularly in low and middle-income countries and in settings where GBV is more culturally embedded. Contextually-appropriate evidence is important to adapt and refine programs to the specific implementation setting. By understanding the effect of training health providers on GBV, policymakers and practitioners in Timor-Leste can better plan and allocate resources toward the health system response to GBV.

### Program Description
This program was a part of the larger Harmonia Activity, a three-year USAID-funded project to address GBV in Timor-Leste. The project was implemented by a local NGO, HAMNASA, in partnership with the Ministry of Health, Ministry of Social Solidarity and Inclusion, and the National Institute of Health. The Harmonia Activity focused on two municipalities within Timor-Leste, Ermera and Liquica, both of which were selected due to their high rates of GBV. When compared to the rest of the country, Ermera and Liquica municipalities have some of the highest rates of GBV, with 58% of women aged 15-49 in Ermera and 51% of women aged 15-49 in Liquica ever experiencing physical, sexual, or emotional violence at the hands of a partner.3

The Harmonia Activity had two components: a community microplanning intervention (a community-based approach to changing patriarchal norms through a Plan-Do-Study-Act model) and implementation of a ‘Responding to GBV’ curriculum for health providers. This paper focuses on the ‘Responding to GBV’ arm of the Harmonia Activity. The goal of the ‘Responding to GBV’ curriculum was to educate health providers in GBV dynamics and shift social norms with the future intent of providing better protection for GBV victims. The curriculum was based on a further adaptation of the WHO curriculum, ‘Caring for women subjected to violence,’  for training health providers on identifying women experiencing violence and providing them essential, women-centered care and access to local resources.12 The curriculum emphasizes the LIVES approach (Listen, Inquire, Validate, Enhance safety and support) in providing care to women who experience violence.12

Facilitators from HAMNASA and the National Institute of Health (INS) were prepared to lead the ‘Responding to GBV’ curriculum via a five-day intensive training-of-trainers (ToT). The ToT was led by trainers from the National University and INS who are experienced at conducting ToTs and are familiar with the training curriculum.

The training intervention took place at regional health facilities and consisted of an initial five days of intensive training, followed by monthly ‘learning lab’ sessions to be carried out at health facilities over a 14-month period. In an effort to create a health-system wide response to GBV, HAMNASA staff worked with INS and the Ministry of Health to recruit participants. Advocacy meetings were conducted in each municipality to increase buy-in from health managers. All health providers in the selected municipalities were eligible to participate in the trainings. Participants included medical doctors or nurses, as well as health facility managers, pharmacists, lab analysts, and other positions. From July to November 2021, eleven intensive trainings took place at health facilities in Ermera and Liquica, training a total of 302 healthcare workers on GBV. These five-day intensive trainings consisted of 14 two-hour modules on various GBV topics. Learning labs, which were shorter 2-hour trainings covering one GBV module each, ran for 14-months following the intensive trainings, from March 2022 to June 2023.

## Aims

Aim 1: To determine if there is significant differences in health provider knowledge, attitudes, empathy, system support, confidence, and practices in responding to GBV before and after participation in the 'Responding to GBV' curriculum in the intervention municipalities in Timor-Leste, using baseline (pre-training) and 14-month post-learning lab data.

Sub-aim 1.1: To evaluate changes in health provider knowledge, attitudes, empathy, system support, confidence, and practices in responding to GBV after participation in the 'Responding to GBV' intensive training compared to the 14-months of follow-up learning labs after the training in the intervention municipalities in Timor-Leste, using percent change from baseline to post-intensive training, and post-intensive to endline.

Aim 2: To identify provider-level characteristics associated with change in GBV knowledge, attitudes, empathy, system support, confidence, and practice scores, including age, sex, occupation, years of professional experience, and attendance at learning labs.

## Methods

### Data Collection
We utilized data collected during the ‘Responding to GBV’ training sessions and follow-up learning labs. To compare the effectiveness of the five-day intensive training and the monthly learning labs in improving health provider competencies to respond to GBV in the selected municipalities of Timor-Leste, the same evaluation tool was used at baseline (prior to training), following the 5- day intensive training, and again at the 14-month endline of the monthly learning labs. The survey was translated from English into Tetum by HAMNASA staff, and then translated back into English to ensure translation accuracy. Participants consented to their responses being used for research purposes on each questionnaire. Participants who did not consent to having their survey responses used for research purposes were excluded from the analysis.

An adaptation of the WHO curriculum evaluation tool was used to assess knowledge, attitudes, confidence, and system support. This evaluation measure consisted of 89 questions, organized into four domains related to the GBV training curriculum: knowledge, attitudes, confidence, and system support. The survey and its related domains can be viewed in the supplemental materials.

The knowledge domain included 43 questions, each assessed with responses of yes, no, or I don’t know, for a maximum possible score of 43. Examples of questions included, “there are common injury patterns associated with domestic violence,” “it is a healthcare provider’s legal duty to help the woman subjected to violence to report it to the police,” and “is depression, anxiety, or chronic stress a warning sign that a woman may have been subjected to domestic or sexual violence?”. Questions were scored as 1 for correct (“yes”) responses and 0 for incorrect (“no” or “I don’t know”) responses. Higher scores were indicative of greater GBV-related knowledge.

The attitudes domain included 29 questions measured on 4 and 5-point Likert scales for a maximum possible score of 102. Examples of questions included “domestic violence is a private matter and people outside the family should not interfere”, “it is humiliating to patients to question them about abuse”, and “it is acceptable for a man to hit his wife or girlfriend if she argues with him.” Questions were formulated in both positive and negative terms. For questions using the 5-point Likert scale, responses were scored on a scale where positively worded items (where “strongly agree” was the preferred answer) were scored as follows: strongly disagree = 0; disagree = 1; neither agree nor disagree = 2; agree = 3; strongly agree = 4. Conversely, negatively worded items (where “strongly disagree” was the desired response) had their scores reversed to account for the preferred response, with strongly disagree obtaining the highest score of 4. For questions using the 4-point Likert scale, the desirable response for all questions was “no, it is never acceptable.” In this case, responses were scored as follows: yes, it is acceptable = 0; sometimes it is acceptable or I don’t know = 1; no, it is never acceptable = 2. Higher scores were indicative of more desirable health provider attitudes toward GBV and the health provider role in responding to GBV.

The confidence domain included 10 questions measured on a 5-point Likert scale for a maximum possible score of 40. Examples of questions included asking health providers if they felt prepared to, “identify a woman who has been subjected to violence by signs and symptoms she reports”, “offer supportive statements to a women subjected to domestic violence or sexual assault”, and “to assess the immediate level of danger for a woman or child after sexual assault or domestic violence.” Responses were scored as follows: not at all prepared = 0; slightly prepared = 1; somewhat prepared = 2; sufficiently prepared = 3; and quite well prepared = 4. Higher scores were indicative of greater confidence in identifying and caring for women subjected to violence.

The system support domain included 6 questions, each assessed with responses of yes, no, or I don’t know, for a maximum possible score of 6. Examples of questions included, “I can readily look up information (e.g. either a guide or standard operating procedure) on how to manage cases of domestic violence or sexual assault” and “my supervisor supports me pro-actively asking my patients or clients about whether they are experiencing domestic violence.” Questions were scored as 1 for correct (“yes”) responses and 0 for incorrect (“no” or “I don’t know”) responses. Higher scores were indicative of a perception of greater resources and support for health providers to respond to GBV within their facility.

The Toronto Empathy Questionnaire (TEQ) was used to provide a measure of empathy for GBV survivors. The TEQ is a brief, unidimensional tool to measure empathy as a broad construct across populations.13 The tool was developed and validated in Canada, where studies demonstrated high internal consistency and reliability amongst university students.13 Subsequently, the TEQ have been administered to assess empathy in various settings, including, but not limited to, medical students in South Korea14 and Romania15, and amongst the general population in the Czech Republic.16 In these settings, the TEQ was translated into the appropriate language and adapted versions of the TEQ were shown to be more valid.14–16 Adaptations included rewording of statements and/or exclusion of survey items.13–15 To date, there is no evidence to support the TEQ’s previous use in Timor-Leste or translation into Tetum.

The empathy domain, derived from the TEQ, included 16 questions measured on a 5-point Likert scale for a total possible score of 64. Examples of questions included, “I enjoy making others feel better”, “I become irritated when someone cries”, and “it upsets me to see someone being treated disrespectfully.” Responses were scored on a scale where positively worded items (where “always” was the preferred answer) were scored as follows: never = 0; rarely = 1; sometimes = 2; often = 3; always = 4. Conversely, negatively worded items (where “never” was the desired response) had the scores reversed to account for the preferred response, with “never” obtaining the highest score of 4. Higher scores were indicative of greater self-reported empathic behavior.

Composite variables for each domain will be created by summing the scores of each survey item within their respective domain. Raw composite scores will be rescaled to reflect the topic area-specific possible points achieved proportional to 100. For instance, a composite score of 4 on a section with 5 possible points would be rescaled to a score of 80 (4/5 * 100) and can be interpreted as a percentage out of 100.

The survey asked two additional questions on provider practices which will be analyzed and reported separately due to it not being applicable to all respondents. The first question asked, “in the past month, have you identified anyone suffering from domestic violence among your female patients? If yes, approximately how many women?” with possible answers of yes, no, and N/A (not in clinical practice during this period) and a space to fill in the number of women identified. The second question used skip logic and was only applicable to those who answered “yes” to the first question. It asked respondents which actions they had taken after identifying a woman subjected to domestic violence, including practices like “offering validating and supportive statements,” “discussing the options she may have,” and “referring the woman to support services available within the community.” Questions were scored as 1 for correct (“yes”) responses and 0 for incorrect (“no” or “I don’t know”) responses. Higher scores were indicative of more appropriate actions taken to support women identified as victims of GBV.

All participants filled out paper survey forms, and the data were subsequently entered into the RedCap data management software by HAMNASA staff. Fewer than 1% of responses were missing from the scored domains.

### Ethical Approval
Written consent was obtained from all participants in the study. The INS Human Research Ethics Committee in Timor-Leste approved this project (ID: 417). The University of Washington Institutional Review Board granted this evaluation a non-research determination (ID: STUDY00018231). 

### Statistical analysis - Need to continue updating 

Only those with survey data for all three timepoints (baseline, post-intensive, and 14-month post-learning lab) were included in the analysis, reducing our sample size while increasing the study’s robustness. Missing demographics were forward filled amongst matched individuals. We used descriptive statistics to describe the characteristics of the study population, including demographic variables such as age group, sex, and occupation. We compared the characteristics of those who remain in the study to those lost to follow-up using a chi-square test of independence to assess potential differences in demographics between the two groups. A t-test will assess differences in baseline scores between the retained and lost to follow-up participants.
To assess change in individual health provider knowledge, attitudes, empathy, confidence, and practices in responding to GBV before, during, and after participation in the 'Responding to GBV' curriculum in the intervention municipalities in Timor-Leste (aim 1), generalized estimating equations regression will be used. GEE is designed to analyze correlated data, making it suitable for analyzing individually matched survey responses over three timepoints. Follow-up scores from each domain at 5 days and 14-months will serve as outcomes and baseline score as the primary independent variable.  Multiple imputation was utilized to complete missing survey responses.  Analysis will be adjusted for age, sex, occupation, and years of professional experience.
 
Secondary analysis will assess the relationship between demographic variables (such as age, sex, occupation, and years of professional experience) and change in knowledge, attitudes, confidence, system support, empathy, and practices over the three timepoints.
 
GEE models will be fit for each of the six survey domains (knowledge, attitudes, empathy, system support, confidence, and practices). We will model the outcome of interest, follow-up score, using a log link function, a Gaussian response distribution, and unstructured correlation. The predictor variables are age group, sex, occupation, and years of professional experience.
 
The following equation will be used to model the outcome of interest follow-up scores:
g(E[Yij]) = a + f31Xij1 + f32Xij2 + f33Xij3 + f34Xij4 + f35Xij5
 
Where:

Yij = follow-up score
a = baseline score
1 = age group
2 = sex
3 = occupation
5 = years of professional experience

-Bias?

All statistical analyses will be conducted using appropriate software such as R or Stata, and statistical significance will be set at p < 0.05.


### Author Positionality
Before presenting the results of this research, it is important to acknowledge my positionality within this research. I am a white American woman pursuing graduate studies in global health. While I have visited Timor-Leste for business, I have not spent substantial time there nor do I speak Tetun. I have experience in conducting healthcare provider training evaluations, but limited experience in gender-based violence research specifically. I have never been a victim of gender-based violence, and believe firmly in its unacceptability. I have benefitted from the mentorship of the HAMNASA team, who have expertise in the cultural context of Timor-Leste and the specifics of the intervention, and my thesis committee, who have expertise in intervention design, research methods, and gender-based violence. 

# Results

## Demographic information

```{r}
demographic_table <- demographic_table %>%
  modify_caption("**Table 1. Participant Characteristics**")
demographic_table
```
Matched data was obtained for 78 individuals, representing a total of 234 surveys out of the 676 conducted across the three timepoints. There were no significant demographic differences between individuals with less than three timepoints and all three timepoints, leading us to believe the smaller longitudinal sample remained representative of the intervention population. Demographic characteristics and work history of participants at the baseline are presented in Table 1. The overall sample had `r inline_text(demographic_table, variable = sex_factored, level = "Female", column= "stat_0")` female and `r inline_text(demographic_table, variable = sex_factored, level = "Male", column= "stat_0")` male participants. `r inline_text(demographic_table, variable = age_groups, level = "25-34 years old", column= "stat_0")` of the sample fell within the 25-34 age group. The most prevalent occupational role was medical doctor, accounting for `r inline_text(demographic_table, variable = position_groups, level = "Medical doctor", column= "stat_0")`, followed by nurses at `r inline_text(demographic_table, variable = position_groups, level = "Nurse", column= "stat_0")`, and midwives at `r inline_text(demographic_table, variable = position_groups, level = "Midwife", column= "stat_0")`. On average, participants had been practicing in their clinical roles for approximately `r inline_text(demographic_table, variable = position_years_clean, column= "stat_0")` years. Only `r inline_text(demographic_table, variable = previous_training_factored, column= "stat_0")` of participants had received any previous GBV training.

## Scores

```{r}
comparison_tbl_1 <- comparison_tbl_1 %>%
  modify_caption("**Table 2. Median participant scores at baseline and endline, by domain**")
comparison_tbl_1
```

Significant improvement in participant scores were observed from baseline (timepoint 1) to endline (timepoint 3) in each domain, as illustrated in table 2 (p <.001). From baseline to endline, median GBV knowledge improved from `r inline_text(comparison_tbl_1, variable = knowledge_overall, column= "stat_1")` to `r inline_text(comparison_tbl_1, variable = knowledge_overall, column= "stat_2")`, out of 43 possible points. The attitudes toward GBV domain improved from `r inline_text(comparison_tbl_1, variable = attitude_overall, column= "stat_1")` to `r inline_text(comparison_tbl_1, variable = attitude_overall, column= "stat_2")`, out of 102 possible points. Confidence scores that represented provider confidence in identifying and treating women subjected to GBV improved from `r inline_text(comparison_tbl_1, variable = confidence_score, column= "stat_1")` to `r inline_text(comparison_tbl_1, variable = confidence_score, column= "stat_2")`, out of 40 possible points. Empathy scores, representing provider's empathic behavior, improved from `r inline_text(comparison_tbl_1, variable = empathy_score, column= "stat_1")` to `r inline_text(comparison_tbl_1, variable = empathy_score, column= "stat_2")`, out of 64 possible points. System support scores rose from `r inline_text(comparison_tbl_1, variable = system_support_score, column= "stat_1")` to `r inline_text(comparison_tbl_1, variable = system_support_score, column= "stat_2")`, out of 6 possible points. While comparatively the practice domain had a small number of respondents (n = `r inline_text(comparison_tbl_1, variable = practice_score, column= "n")`) due to the skip logic employed in the practice questions, the results were still significant. Practice scores, measuring appropriate actions taken by providers to support women identified as victims of GBV, improved from `r inline_text(comparison_tbl_1, variable = practice_score, column= "stat_1")` to `r inline_text(comparison_tbl_1, variable = practice_score, column= "stat_2")`.

While significant improvements were seen from intervention baseline to endline, scores between timepoint 2 (post 5-day intensive GBV training) and timepoint 3 (endline, following 14-months of monthly learning labs) plateaued (figure 1).

## Regression results