---
title: "Monitoring, Evalution and Learning Report"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
   bookdown::word_document2:
      fig_caption: yes
      toc: true
      toc_depth: 2
      reference_docx: "../template/template.docx"
---

```{r, setup, include=FALSE, warning = FALSE, message=FALSE}
library(stringr)
library(officedown)
library(officer) 
library(knitr)
current_wd <- getwd()

if(endsWith(current_wd, "gbv-health-provider-study")){
  disc_project_wd <- current_wd
} else if (endsWith(current_wd, "/paper")) {
  disc_project_wd <- str_remove(current_wd, "/paper")   
} else {
  print("Got a WD that's not handled in the If-else ladder yet")
}

knitr::opts_chunk$set(fig.align = 'center',
                      echo = FALSE,
                      warning = FALSE, 
                      message = FALSE,
                      dpi = 300)

## SOURCE FIGURES AND TABLES
source(paste(disc_project_wd, "/code/table_1.R", sep = ""))
source(paste(disc_project_wd, "/code/mel_report_analysis.R", sep = ""))
source(paste(disc_project_wd, "/code/figures.R", sep = ""))
```

# Background
As a part of the three-year USAID-funded Harmonia Activity, HAMNASA partnered with the Timor-Leste Ministry of Health, Ministry of Social Solidarity and Inclusion, and the National Institute of Health to address gender-based violence (GBV) through engaging healthcare providers to shift social norms and protect victims of violence. The Harmonia activity focused on two municipalities within Timor-Leste, Ermera and Liquica, both which experience high rates of GBV.

From July to November 2021, eleven intensive trainings took place at health facilities in Ermera and Liquica, training a total of 232 healthcare workers on GBV. These five-day intensive trainings consisted of 14 two-hour modules on various gender-based violence topics. Learning labs, which were shorter 2-hour trainings covering one GBV module each, ran for 14-months, from March 2022 to June 2023.

## Monitoring and Evaluation Targets
Three indicators included in the Harmonia Activity’s monitoring and evaluation plan will be used to assess the success of the ‘Responding to GBV Learning Labs’ Curriculum (Table 1). Two are direct measures of the success of the ‘Responding to GBV Learning Labs’ Curriculum (outcomes 4 and 5). The third measure (outcome 7) will be reported as a complementary finding to the official monitoring and evaluation plan’s use of health facility data to identify change in the number of clients being identified as victims of GBV.

| Result                                                                                                                | Indicator                                                                               | Frequency of Collection                                                                                                   | Unit of Measure                                                                                                                                         | Value          |
|-----------------------------------------------------------------------------------------------------------------------|-----------------------------------------------------------------------------------------|---------------------------------------------------------------------------------------------------------------------------|----------------------------------------------------------------------------------------------------------------------------------------------------------|----------------|
| Outcome 4: Health providers who have participated in the ‘Responding to GBV Learning Lab’ Curriculum demonstrate improved knowledge related to GBV | Indicator 4: % of health facilities demonstrate improvement in staff knowledge, attitude, and empathy in responding to GBV | Three time points: pre-training, post-intensive training, and post-intervention                                             | Average composite score of knowledge, attitude, and empathy domains by health facility cohort pre- to post-intervention | 80% of health facilities |
| Outcome 5: Health providers who have participated in the ‘Responding to GBV Learning Lab’ have increased confidence in delivering survivor-centered care   | Indicator 5: % of health facilities that demonstrate improvement in health provider confidence in responding to GBV | Three time points: pre-training, post-intensive training, and post-intervention                                             | Average composite score of the confidence, system support, and professional role domains  by health facility cohort pre to post-intervention | 80% of health facilities |
| Outcome 7: Increase in clients identified who are victims of GBV                                                     | Indicator 7: % increase in GBV clients identified                                         | Two timepoints: pre-training and post-intervention* *This measure can be used to compare to health facility reporting (which is the official MEL data source) on clients identified | Percentage compared to baseline | 50% increase   |

# Methods
## Data Collection
A pre- and post-survey was adapted from the WHO curriculum evaluation tool, with the Toronto Empathy Questionnaire (TEQ) added. The 105 question survey was translated from English into Tetum by HAMNASA staff and was administered immediately pre- and post-intensive training, as well as at an endline survey administered following the monthly learning labs in June and July 2023. 

All surveys were completed on paper forms and entered into RedCap data management software by HAMNASA staff. Participants were assigned unique identifiers that were utilized in combination with facility, sex, and position to match individuals across the three timepoints. Only individuals who completed all three surveys (pre-training, post-intensive, and post-intervention) were included in the analysis in order to measure change over time.

## Data Analysis
<u>Pre-and post-training scores by topic area</u>
The evaluation tool included 11 topic areas across 5 domains (knowledge, attitudes, confidence, empathy, and system support). Two domains, knowledge and attitudes, each had 4 sub-domains. Scores for knowledge and attitudes are reported by sub-domain to increase specificity. 

True/false and yes/no questions were scored 1 for correct answers and 0 for incorrect, unanswered questions, or responses of “I don’t know.” Questions using a Likert scale were scored as the number of responses minus 1, with the least desirable response being scored as 0. 

For the empathy and attitude domains only, questions were worded both positively and negatively. Responses were scored according to the following scale for positively worded items (for which preferred answer was “always”): never = 0; rarely = 1; sometimes = 2; often = 3; always = 4. For negatively worded items (for which preferred answer was “never”), the scores were reversed to account for the desired response.

Composite variables for each domain and subdomain were created by summing the scores of each survey item within each the respective domain. All domains’ scores were normalized on a 100-point scale. 

<u>Self-reported practices</u>
Participants self-reported practices in identifying and caring for women subjected to violence in the survey. There should not have been variation in practices from pre- to post-intensive, so only the baseline (pre-survey) measure was used from the intensive. The proportion of participants who identified a woman suffering from domestic violence among their female patients in the past month was calculated. Among those providers who identified women suffering from violence, the average number of patients identified and the proportion of appropriate actions taken were calculated and compared at baseline (pre-intensive training) to endline (post-learning labs).

# Results
## Demographic information
Matched data was available for 46 individuals (138 surveys) from a total of 676 surveys over the three timepoints. Participant demographic characteristics and work history at baseline are presented in Table 2.

The overall sample was 63% female (29 participants) and 37% male (17 participants). 50% of the sample was between 25-34 years old (22 participants). The most common position held was medical doctor (32% or 14 participants), followed by nurse (27% or 12 participants) and midwife (25% or 11 participants). At baseline, participants had been practicing in their clinical roles for an average of 7 years. Most participants saw an average of 39 or fewer patients per week, with 27% (12 participants) seeing less than 20 patients per week and another 27% seeing between 20-39 patients per week. 24% of participants (11 participants) saw 60 or more patients per week. Only 8.7% of participants (4 participants) had ever received GBV training before.

```{r}
demographic_table %>%
  modify_caption("**Table 1. Participant Characteristics at Baseline**")
```

## Outcome 4: Health providers who have participated in the ‘Responding to GBV Learning Lab’ Curriculum demonstrate improved knowledge related to GBV.
When comparing the results to indicator 4, which is the percentage of health facilities demonstrate improvement in staff knowledge, attitude, and empathy in responding to GBV, the project did not meet/met/exceeded. ____% of  intervention facilities demonstrated improvement in staff knowledge, attitudes, and empathy in responding to GBV (table 2). Additional analysis of this indicator can be found in the Appendix.

```{r}
outcome4_table <- difference_by_facility %>%
  select(region, outcome4_mean_value_1, outcome4_mean_value_3, outcome4_difference) %>%
flextable() 

outcome4_table <- set_header_labels(
  table,
  region = "Region",
  outcome4_mean_value_1 = "Mean score timepoint 1", 
  outcome4_mean_value_3 = "Mean score timepoint 3", 
  outcome4_difference = "Difference in mean scores") 

outcome4_table
```

## Outcome 5: Health providers who have participated in the ‘Responding to GBV Learning Lab’ have increased confidence in delivering survivor-centered care.
When comparing the results to indicator 5, which is the percentage of health facilities that demonstrate improvement in health provider confidence in responding to GBV, the project did not meet/met/exceeded. ____% of  intervention facilities demonstrated improvement in staff confidence in responding to GBV (table 3). Additional analysis of this indicator can be found in the Appendix.

## Outcome 7: Increase in clients identified who are victims of GBV.
At baseline, healthcare providers (n = 6) self-reported identifying an average of 1.5 clients who were victims of GBV in the past month. At endline, healthcare providers (n = 16) self-reported identifying an average of 3 clients who were victims of GBV in the past month. These self-reported numbers exceeded indicator seven’s target of a 50% increase in clients identified. Per self-reports, the number of providers identifying clients experiencing GBV within the past month increased by 167% and the number of clients identified increased by 100%. Facility records will be assessed to capture a more holistic picture of client identification in the facility. 

Of those healthcare providers who identified clients within the past month who were victims of GBV, there was an improvement in the actions taken from baseline to endline (table __).

## Learning Labs
While the above scores demonstrate the impact of the entirety of the ‘Responding to GBV Learning Labs’ Curriculum, inclusive of the 5-day intensive training and the 14-month Learning Lab intervention, we were interested in assessing what occurred between the end of the 5-day intensive training (the post-intensive assessment) and the end of the learning labs (endline). 


# Interpretation 

# Appendix 
## All participant scores
```{r}
figure_fp <- paste(gbv_project_wd, "/figures/mean_scores_bar_chart.png", sep = "")

knitr::include_graphics(figure_fp)

```
