---
title: "Monitoring, Evalution and Learning Report"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
   bookdown::word_document2:
      fig_caption: yes
      toc: true
      toc_depth: 2
      reference_docx: "../template/template.docx"
---

```{r, setup, include=FALSE, warning = FALSE, message=FALSE}
options(repos = "https://cloud.r-project.org/")

library(stringr)
library(officedown)
library(officer) 
library(knitr)
library(gtsummary)
current_wd <- getwd()

if(endsWith(current_wd, "gbv-health-provider-study")){
  disc_project_wd <- current_wd
} else if (endsWith(current_wd, "/paper")) {
  disc_project_wd <- str_remove(current_wd, "/paper")   
} else {
  print("Got a WD that's not handled in the If-else ladder yet")
}

knitr::opts_chunk$set(fig.align = 'center',
                      echo = FALSE,
                      warning = FALSE, 
                      message = FALSE,
                      dpi = 300)

## SOURCE FIGURES AND TABLES
source(paste(disc_project_wd, "/code/participant_id_cleaning.R", sep = ""))
source(paste(disc_project_wd, "/code/table_1.R", sep = ""))
source(paste(disc_project_wd, "/code/mel_report_analysis.R", sep = ""))
source(paste(disc_project_wd, "/code/figures.R", sep = ""))
source(paste(disc_project_wd, "/code/table_scores.R", sep = ""))
source(paste(disc_project_wd, "/code/indicator_table.R", sep = ""))
```

# Background
As a part of the three-year USAID-funded Harmonia Activity, HAMNASA partnered with the Ministry of Health, Ministry of Social Solidarity and Inclusion, and the National Institute of Health to address gender-based violence (GBV) in Timor-Leste. The Harmonia activity focused on two municipalities within Timor-Leste, Ermera and Liquica, both of which experience high rates of GBV. The Harmonia activity had two components: a community microplanning intervention and implementation of a 'Responding to GBV' curriculum for healthcare providers. This report aims to evaluate the overall impact of the 'Responding to GBV' curriculum. 

The 'Responding to GBV' curriculum's goal was to engage healthcare providers to shift social norms and protect victims of violence. The curriculum was based upon the WHO curriculum for training healthcare providers caring for women subjected to GBV and its previous adaptation for nursing, midwifery and medical degrees in Timor-Leste.

The curriculum included an initial five days of intensive training, followed by monthly ‘learning lab’ sessions to be carried out at health facilities over a 14-month period. In an effort to create a health-system wide response to GBV, participants included medical doctors or nurses, as well as health facility managers, pharmacists, lab analysts, and other positions. From July to November 2021, eleven intensive trainings took place at health facilities in Ermera and Liquica, training a total of 232 healthcare workers on GBV. These five-day intensive trainings consisted of 14 two-hour modules on various gender-based violence topics. Learning labs, which were shorter 2-hour trainings covering one GBV module each, ran for 14-months following the intensive trainings, from March 2022 to June 2023.

## Monitoring, Evaluation, and Learning Targets
Three indicators included in the Harmonia Activity’s monitoring, evaluation, and learning plan will be used to assess the success of the ‘Responding to GBV Learning Labs’ Curriculum (Table 1). Two are direct measures of the success of the ‘Responding to GBV Learning Labs’ Curriculum (outcomes 4 and 5). The third measure (outcome 7) will be reported as a complementary finding to the official monitoring, evaluation, and learning plan’s use of health facility data to identify change in the number of clients being identified as victims of GBV.

<!-- | Result                                                                                                                | Indicator                                                                               | Frequency of Collection                                                                                                   | Unit of Measure                                                                                                                                         | Value          | -->
<!-- |-----------------------------------------------------------------------------------------------------------------------|-----------------------------------------------------------------------------------------|---------------------------------------------------------------------------------------------------------------------------|----------------------------------------------------------------------------------------------------------------------------------------------------------|----------------| -->
<!-- | Outcome 4: Health providers who have participated in the ‘Responding to GBV Learning Lab’ Curriculum demonstrate improved knowledge related to GBV | Indicator 4: % of health facilities demonstrate improvement in staff knowledge, attitude, and empathy in responding to GBV | Three time points: pre-training, post-intensive training, and post-intervention                                             | Average composite score of knowledge, attitude, and empathy domains by health facility cohort pre- to post-intervention | 80% of health facilities | -->
<!-- | Outcome 5: Health providers who have participated in the ‘Responding to GBV Learning Lab’ have increased confidence in delivering survivor-centered care   | Indicator 5: % of health facilities that demonstrate improvement in health provider confidence in responding to GBV | Three time points: pre-training, post-intensive training, and post-intervention                                             | Average composite score of the confidence, system support, and professional role domains  by health facility cohort pre to post-intervention | 80% of health facilities | -->
<!-- | Outcome 7: Increase in clients identified who are victims of GBV                                                     | Indicator 7: % increase in GBV clients identified                                         | Two timepoints: pre-training and post-intervention* *This measure can be used to compare to health facility reporting (which is the official MEL data source) on clients identified | Percentage compared to baseline | 50% increase   | -->

```{r}
indicator_table
```

# Methods
## Data Collection
A pre- and post-survey was adapted from the WHO curriculum evaluation tool, with the Toronto Empathy Questionnaire (TEQ) added. The 105 question survey was translated from English into Tetum by HAMNASA staff and was administered immediately pre- and post-intensive training, as well as at post-intervention following the monthly learning labs in June and July 2023. 

All surveys were completed on paper forms and entered into RedCap data management software by HAMNASA staff. Participants were assigned unique identifiers that were utilized in combination with facility, sex, and position to match individuals across the three timepoints, pre-training (baseline), post-intensive, and post-intervention (endline). Only individuals who completed all three surveys were included in the analysis in order to measure change over time.

## Data Analysis
**Pre-and post-training scores by topic area**
The evaluation tool included 11 topic areas across 5 domains (knowledge, attitudes, confidence, empathy, and system support). Outcome 4 is a composite of the knowledge, attitude*, and empathy domains. Outcome 5 is a composite of the confidence and system support domains and the attitudes toward professional role sub-domain. 

*The attitude domain has four sub-domains: general attitudes toward GBV, GBV unacceptability, attitudes towards gender roles, and attitudes toward professional role in responding to GBV. The attitudes toward professional role in responding to GBV sub-domain was included in outcome 5 rather than outcome 4.

True/false and yes/no questions were scored 1 for correct answers and 0 for incorrect, unanswered questions, or responses of “I don’t know.” Questions using a Likert scale were scored as the number of responses minus 1, with the least desirable response being scored as 0. 

For the empathy and attitude domains only, questions were worded both positively and negatively. Responses were scored according to the following scale for positively worded items (for which preferred answer was “always”): never = 0; rarely = 1; sometimes = 2; often = 3; always = 4. For negatively worded items (for which preferred answer was “never”), the scores were reversed to account for the desired response.

Composite variables for each domain and subdomain were created by summing the scores of each survey item within each the respective domain. All domains’ scores were normalized on a 100-point scale. 

Given the small sample size, CHCs and health posts were collapsed into one "CHC catchment area."

**Self-reported practices**
Participants self-reported practices in identifying and caring for women subjected to GBV in the survey. The proportion of providers who identified a woman suffering from GBV among their female patients in the past month was calculated. Among those providers who identified women suffering from GBV, the average number of patients identified and the proportion of appropriate actions taken were calculated at baseline and endline.

# Results
## Demographic information
Matched data was available for 46 individuals (138 surveys) from a total of 676 surveys over the three timepoints. Participant demographic characteristics and work history at baseline are presented in Table 2.

The overall sample was 63% female (29 participants) and 37% male (17 participants). 50% of the sample was between 25-34 years old (22 participants). The most common position held was medical doctor (32% or 14 participants), followed by nurse (27% or 12 participants) and midwife (25% or 11 participants). At baseline, participants had been practicing in their clinical roles for an average of 7 years. Most participants saw an average of 39 or fewer patients per week, with 27% (12 participants) seeing less than 20 patients per week and another 27% seeing between 20-39 patients per week. 24% of participants (11 participants) saw 60 or more patients per week. Only 8.7% of participants (4 participants) had ever received GBV training before.

```{r}
demographic_table %>%
  modify_caption("**Table 2. Participant Characteristics at Baseline**")
```

## Outcome 4: Health providers who have participated in the ‘Responding to GBV Learning Lab’ Curriculum demonstrate improved knowledge related to GBV.
When assessing indicator 4, the percentage of health facilities demonstrating improvement in staff knowledge, attitude, and empathy in responding to GBV, the project exceeded the MEL target, with 90% of   facilities demonstrating improvement in staff knowledge, attitudes, and empathy in responding to GBV (table 3). Additional analysis of this indicator can be found in the Appendix.

```{r}
outcome4_table <- difference_by_facility %>%
  select(region, outcome4_mean_value_1, outcome4_mean_value_3, outcome4_difference) %>%
flextable()

outcome4_table <- set_caption(outcome4_table, "**Table 3. Outcome 4: Mean Provider Knowledge Score by Facility Catchment Area**")

outcome4_table <- set_header_labels(
  outcome4_table,
  region = "Catchment area",
  outcome4_mean_value_1 = "Baseline",
  outcome4_mean_value_3 = "Endline",
  outcome4_difference = "Percent Difference")

footer_values <- c("All Facilities", 
                   mean(difference_by_facility$outcome4_mean_value_1, na.rm = TRUE),
                   mean(difference_by_facility$outcome4_mean_value_3, na.rm = TRUE),
                   mean(difference_by_facility$outcome4_difference, na.rm = TRUE))

outcome4_table <- add_footer_row(outcome4_table, values = footer_values, colwidths = c(1,1,1,1))

outcome4_table
```

## Outcome 5: Health providers who have participated in the ‘Responding to GBV Learning Lab’ have increased confidence in delivering survivor-centered care.
When comparing the results to indicator 5, which is the percentage of health facilities that demonstrate improvement in health provider confidence in responding to GBV, the project exceeded the MEL target, with 100% of facilities demonstrating improvement in staff confidence in responding to GBV (table 4). Additional analysis of this indicator can be found in the Appendix.
```{r}
outcome5_table <- difference_by_facility %>%
  select(CHC_catchment, outcome5_mean_value_1, outcome5_mean_value_3, outcome5_difference) %>%
flextable()

outcome5_table <- set_caption(outcome5_table, "**Table 4. Outcome 5: Mean Provider Confidence Score by Facility Catchment Area**")

outcome5_table <- set_header_labels(
  outcome5_table,
  CHC_catchment = "Catchment area",
  outcome5_mean_value_1 = "Baseline",
  outcome5_mean_value_3 = "Endline",
  outcome5_difference = "Percent Difference")
  
footer_values <- c("All Facilities", 
                   mean(difference_by_facility$outcome5_mean_value_1, na.rm = TRUE),
                   mean(difference_by_facility$outcome5_mean_value_3, na.rm = TRUE),
                   mean(difference_by_facility$outcome5_difference, na.rm = TRUE))

outcome5_table <- add_footer_row(outcome5_table, values = footer_values, colwidths = c(1,1,1,1))

outcome5_table
```

## Outcome 7: Increase in clients identified who are victims of GBV.
At baseline, healthcare providers (n = 6) self-reported identifying an average of 1.5 clients who were victims of GBV in the past month. At endline, healthcare providers (n = 16) self-reported identifying an average of 3 clients who were victims of GBV in the past month. These self-reported numbers exceeded indicator seven’s target of a 50% increase in clients identified. Per self-reports, the number of providers identifying clients experiencing GBV within the past month increased by 167% and the number of clients identified increased by 100%.

Of those healthcare providers who identified clients within the past month who were victims of GBV, there was improvement on all but one appropriate practice from baseline to endline (table 5). The one practice without improvement, offering validating and supportive statements, was already at 100% at baseline and therefore could not improve further.

```{r}
pract19_clean_vars <- names(clean_data)[str_detect(names(clean_data), "practices_clean_19")]

outcome7pre <- clean_data %>%
  filter(status == "All three" & time_point == 1) %>%
  select(all_of(pract19_clean_vars)) %>%
  na.omit() %>%
  t() %>%
  as.data.frame() %>%
  mutate(percent_correct_pre = (rowSums(.))/6*100) %>%
  select(percent_correct_pre)
outcome7pre$question <- rownames(outcome7pre)

outcome7post <- clean_data %>%
  filter(status == "All three") %>%
  filter(time_point == 3) %>%
  select(all_of(pract19_clean_vars)) %>%
  na.omit() %>%
  t() %>%
  as.data.frame() %>%
  mutate(percent_correct_post = (rowSums(.))/16*100) %>%
  select(percent_correct_post)
outcome7post$question <- rownames(outcome7post)

outcome7_pre_post <- merge(outcome7pre, outcome7post, by = "question", all = TRUE) %>%
  mutate(difference = percent_correct_post - percent_correct_pre) %>% 
  mutate(question = factor(question, levels = c("practices_clean_19a", 
                                                        "practices_clean_19b",
                                                        "practices_clean_19c",
                                                        "practices_clean_19d",
                                                        "practices_clean_19e",
                                                        "practices_clean_19f",
                                                        "practices_clean_19g",
                                                        "practices_clean_19h",
                                                        "practices_clean_19i"),
                              labels = c("Assessed the immediate level of danger for the woman",
                    "Discussed the options she may have",
                    "Documented domestic violence history and physical examination findings",
                    "Helped the woman to create a plan to increase her and her children’s safety",
                    "Offered validating and supportive statements",
                    "Provided basic information about domestic violence",
                    "Provided education or resource materials about domestic violence",
                    "Referred the woman to support services in the community",
                    "Talked to the woman about her needs")))

numeric_columns <- sapply(outcome7_pre_post, is.numeric)
outcome7_pre_post[numeric_columns] <- round(outcome7_pre_post[numeric_columns], digits = 2)
  
outcome7_table <- flextable(outcome7_pre_post)

outcome7_table <- set_header_labels(
  outcome7_table,
  question = "Action Taken",
  percent_correct_pre = "Baseline",
  percent_correct_post = "Endline",
  difference = "Percent difference")

outcome7_table <- set_caption(outcome7_table, "**Table 5. Outcome 7: Provider Practices in Identifying and Caring for Victims of GBV**")

outcome7_table
```

## Learning Labs
While the above analysis demonstrates the impact of the entirety of the ‘Responding to GBV Learning Labs’ Curriculum, inclusive of the 5-day intensive training and the 14-month Learning Lab intervention, additional analysis was conducted to understand trends between the end of the 5-day intensive training (the post-intensive assessment) and endline (table 6).
```{r}
clean_scores <- clean_scores %>%
  mutate(timepoint_factored = factor(time_point,
                                     levels = c(2, 3),
                                     labels = c("Post-intensive Training", "Endline")))

learning_labs_table <-
  clean_scores %>%
  filter(status == "All three") %>%  
  filter(time_point != 1) %>%
  tbl_summary(
    include = c("outcome4_score", "outcome5_score", "timepoint_factored"), 
    by = c(timepoint_factored),
    type = c(
      outcome4_score, outcome5_score
    ) ~ "continuous",
    label = list(
      outcome4_score ~ "Outcome 4 Score",
      outcome5_score ~ "Outcome 5 Score"
    ),
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
    ),
    digits = all_continuous() ~ 2
  ) %>%
  add_p() %>%
  bold_p()

learning_labs_table <- learning_labs_table %>%
  modify_caption("**Table 6. Outcome 4 & 5 Scores from Post-intensive Training to Endline**")

learning_labs_table
```

# Conclusion 
There was improvement on all MEL indicators (outcomes 4, 5, and 7) from baseline to endline. These results point to the 'Responding to GBV' curriculum's ability to impact healthcare providers' GBV knowledge and confidence in providing care to victims of GBV. This is particularly important given the high levels of GBV in the intervention municipalities and the low level of previous GBV training participants had received. 

While there was an overall improvement in knowledge and confidence scores, there was large variation in individual facilities' improvement, including a decrease in outcome 4's scores at endline for CHC Ermera. It may be worthwhile to investigate facility characteristics in the future to determine what may be causing the variation in improvement, and characteristics associated with more improvement.

While the number of providers who identified a woman who was a victim of GBV in the past month was relatively small, there were substantial improvements in both the number of providers who identified a woman who was a victim of GBV in the past month and the use of appropriate practices in caring for women who are victims of GBV. The largest changes from baseline to endline in the use of appropriate practices were: referring the woman to support services in the community, helping the woman to create a plan to increase her and her children’s safety, and talking to the woman about her needs. Six of the nine appropriate practices in caring for a woman experiencing GBV were reportedly used by 100% of providers at endline. This is an important finding as it indicates that improvements in GBV knowledge and confidence appear to translate to changes in clinical practice.

While this report primarily looked at changes from baseline to endline, there was interest in understanding the impact of the learning labs on provider knowledge and confidence. Outcomes 4 and 5 both improved, though outcome 4 was not significant, after the intensive training. This may indicate that the learning labs maintained existing knowledge and continued to improve knowledge following the intensive training. In the absence of a counterfactual, however, it is not possible to say with certainty what would have happened without the learning labs. Investigating the utility of the learning labs in addition to the intensive trainings provides an opportunity for future research. [insert evidence about knowledge drop-off in other training programs]

# Appendix 
## Participant Scores without Facility Categories
```{r}
figure_fp <- paste(gbv_project_wd, "/figures/mean_scores_bar_chart.png", sep = "")

knitr::include_graphics(figure_fp)

```
