---
title: "Monitoring, Evalution and Learning Report"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
   bookdown::word_document2:
      fig_caption: yes
      toc: true
      toc_depth: 2
      reference_docx: "../template/template.docx"
---

```{r, setup, include=FALSE, warning = FALSE, message=FALSE}
options(repos = "https://cloud.r-project.org/")

library(stringr)
library(officedown)
library(officer) 
library(knitr)
library(gtsummary)
current_wd <- getwd()

if(endsWith(current_wd, "gbv-health-provider-study")){
  disc_project_wd <- current_wd
} else if (endsWith(current_wd, "/paper")) {
  disc_project_wd <- str_remove(current_wd, "/paper")   
} else {
  print("Got a WD that's not handled in the If-else ladder yet")
}

knitr::opts_chunk$set(fig.align = 'center',
                      echo = FALSE,
                      warning = FALSE, 
                      message = FALSE,
                      dpi = 300)

## SOURCE FIGURES AND TABLES
source(paste(disc_project_wd, "/code/participant_id_cleaning.R", sep = ""))
source(paste(disc_project_wd, "/code/table_1.R", sep = ""))
source(paste(disc_project_wd, "/code/mel_report_analysis.R", sep = ""))
source(paste(disc_project_wd, "/code/figures.R", sep = ""))
source(paste(disc_project_wd, "/code/table_scores.R", sep = ""))
source(paste(disc_project_wd, "/code/indicator_table.R", sep = ""))
```

# Background
As a part of the three-year USAID-funded Harmonia Activity, HAMNASA partnered with the Ministry of Health, Ministry of Social Solidarity and Inclusion, and the National Institute of Health to address gender-based violence (GBV) in Timor-Leste. The Harmonia activity focused on two municipalities within Timor-Leste, Ermera and Liquica, both of which experience high rates of GBV. The Harmonia activity had two components: a community microplanning intervention and implementation of a 'Responding to GBV' curriculum for healthcare providers. This report aims to evaluate the overall impact of the 'Responding to GBV' curriculum. 

The 'Responding to GBV' curriculum's goal was to engage healthcare providers to shift social norms and protect victims of violence. The curriculum was based upon the WHO curriculum for training healthcare providers caring for women subjected to GBV and its previous adaptation for nursing, midwifery and medical degrees in Timor-Leste.

The curriculum included an initial five days of intensive training, followed by monthly ‘learning lab’ sessions to be carried out at health facilities over a 14-month period. In an effort to create a health-system wide response to GBV, participants included medical doctors or nurses, as well as health facility managers, pharmacists, lab analysts, and other positions. From July to November 2021, eleven intensive trainings took place at health facilities in Ermera and Liquica, training a total of 232 healthcare workers on GBV. These five-day intensive trainings consisted of 14 two-hour modules on various gender-based violence topics. Learning labs, which were shorter 2-hour trainings covering one GBV module each, ran for 14-months following the intensive trainings, from March 2022 to June 2023.

## Monitoring, Evaluation, and Learning Targets
Three indicators included in the Harmonia Activity’s Monitoring, Evaluation, and Learning (MEL) Plan will be used to assess the success of the ‘Responding to GBV Learning Labs’ Curriculum (Table 1). Two are direct measures of the success of the ‘Responding to GBV Learning Labs’ Curriculum (outcomes 4 and 5). The third measure (outcome 7) will be reported as a complementary finding to the MEL Plan's use of health facility data to identify change in the number of clients being identified as victims of GBV.

```{r}
indicator_table
```

# Methods
## Data Collection
A survey consisting of 105 questions, which included the Toronto Empathy Questionnaire (TEQ), was used before and after the training program. The survey was adapted from the WHO curriculum evaluation tool and was translated from English into Tetum by HAMNASA staff. It was administered at three different time points: before the intensive trainings, immediately after the intensive trainings, which took place between July and October 2021, and following the fourteen monthly learning lab sessions in June and July 2023.

All participants filled out paper survey forms, and the data were subsequently entered into the RedCap data management software by HAMNASA staff. To track individuals' progress across the three timepoints (pre-training, post-intensive, and post-intervention), participants were assigned unique identifiers. Only those who completed all three surveys were included in the analysis to assess changes over time.

## Data Analysis
**Pre-and post-training scores by topic area**
The evaluation tool assessed 11 topic areas organized into 5 domains: knowledge, attitudes, confidence, empathy, and system support. Outcome 4 represents a combination of the knowledge, attitude, and empathy domains, while Outcome 5 comprises the confidence and system support domains along with the attitudes toward the professional role sub-domain.

The attitude domain consists of four sub-domains: general attitudes toward GBV, GBV unacceptability, attitudes towards gender roles, and attitudes toward the professional role in responding to GBV. The sub-domain related to attitudes toward the professional role in responding to GBV was included in Outcome 5 rather than Outcome 4.

True/false and yes/no questions were scored as 1 for correct answers and 0 for incorrect responses, unanswered questions, or responses such as "I don't know." Questions using a Likert scale were scored as the number of responses minus 1, with the least favorable response receiving a score of 0.

For the empathy and attitude domains only, questions were formulated in both positive and negative terms. Responses were scored on a scale where positively worded items (where "always" was the preferred answer) were scored as follows: never = 0; rarely = 1; sometimes = 2; often = 3; always = 4. Conversely, negatively worded items (where "never" was the desired response) had their scores reversed to account for the preferred response.

Composite variables for each domain and subdomain were created by summing the scores of each survey item within their respective domain. All domain scores were then normalized on a 100-point scale.

Due to the limited sample size, Community Health Centers (CHCs) and health posts were combined into a single category referred to as the "CHC catchment area."

**Self-Reported Practices**
In the survey, participants self-reported their practices in identifying and caring for women who have experienced GBV. We calculated the proportion of providers who identified women experiencing GBV among their female patients in the past month. Among those providers who identified such cases, we calculated the average number of patients identified and the proportion of appropriate actions taken at both baseline and endline.

# Results
## Demographic information
Matched data was obtained for 46 individuals, representing a total of 138 surveys out of the 676 conducted across the three timepoints. Demographic characteristics and work history of participants at the baseline are presented in Table 2.

The overall sample was r inline_text(demographic_table, variable = sex_factored, level = "Female", column = "stat_0") female and inline_text(demographic_table, variable = sex_factored, level = "Male", column = "stat_0") male. inline_text(demographic_table, variable = age_groups, level = "25-34 years old", column = "stat_0") of the sample fell within the 25-34 age group. The most prevalent occupational role was medical doctor, accounting for inline_text(demographic_table, variable = Position, level = "Medical doctor", column = "stat_0"), followed by nurses at inline_text(demographic_table, variable = Position, level = "Nurse", column = "stat_0"), and midwives at inline_text(demographic_table, variable = Position, level = "Midwife", column = "stat_0"). On average, participants had been practicing in their clinical roles for approximately inline_text(demographic_table, variable = Years of practice, column = "stat_0") years. The majority of participants saw fewer than 39 patients per week, with 27% attending to less than 20 patients weekly and another 27% managing between 20-39 patients per week. In contrast, 24% of participants attended to 60 or more patients per week. Only 8.7% of participants (4 individuals) had received any previous GBV training.

```{r}
demographic_table %>%
  modify_caption("**Table 2. Participant Characteristics at Baseline**")
```

## Outcome 4: Health providers who have participated in the ‘Responding to GBV Learning Lab’ Curriculum demonstrate improved knowledge related to GBV.
When evaluating Outcome 4, which measures the percentage of healthcare facilities showing enhancement in staff knowledge, attitudes, and empathy regarding GBV response, the project surpassed the MEL target. Specifically, 90% of the facilities exhibited improvements in staff knowledge, attitudes, and empathy in dealing with GBV, as detailed in Table 3. Further analysis of this indicator is available in the Appendix.
```{r}
outcome4_table <- difference_by_facility %>%
  select(region, outcome4_mean_value_1, outcome4_mean_value_3, outcome4_difference) %>%
flextable()

outcome4_table <- set_caption(outcome4_table, "**Table 3. Outcome 4: Mean Provider Knowledge Score by Facility Catchment Area**")

outcome4_table <- set_header_labels(
  outcome4_table,
  region = "Catchment area",
  outcome4_mean_value_1 = "Baseline",
  outcome4_mean_value_3 = "Endline",
  outcome4_difference = "Difference in mean scores")

footer_values <- c("All Facilities", 
                   mean(difference_by_facility$outcome4_mean_value_1, na.rm = TRUE),
                   mean(difference_by_facility$outcome4_mean_value_3, na.rm = TRUE),
                   mean(difference_by_facility$outcome4_difference, na.rm = TRUE))

outcome4_table <- add_footer_row(outcome4_table, values = footer_values, colwidths = c(1,1,1,1))

outcome4_table
```

## Outcome 5: Health providers who have participated in the ‘Responding to GBV Learning Lab’ have increased confidence in delivering survivor-centered care.
When evaluating Indicator 5, which gauges the percentage of health facilities showcasing enhanced health provider confidence in GBV response, the project not only met but exceeded the MEL target. Specifically, Table 4 reveals that all facilities, constituting 100%, demonstrated improvements in staff confidence when responding to GBV. Further analysis of this indicator is available in the Appendix.

```{r}
outcome5_table <- difference_by_facility %>%
  select(region, outcome5_mean_value_1, outcome5_mean_value_3, outcome5_difference) %>%
flextable()

outcome5_table <- set_caption(outcome5_table, "**Table 4. Outcome 5: Mean Provider Confidence Score by Facility Catchment Area**")

outcome5_table <- set_header_labels(
  outcome5_table,
  region = "Catchment area",
  outcome5_mean_value_1 = "Baseline",
  outcome5_mean_value_3 = "Endline",
  outcome5_difference = "Difference in mean scores")
  
footer_values <- c("All Facilities", 
                   mean(difference_by_facility$outcome5_mean_value_1, na.rm = TRUE),
                   mean(difference_by_facility$outcome5_mean_value_3, na.rm = TRUE),
                   mean(difference_by_facility$outcome5_difference, na.rm = TRUE))

outcome5_table <- add_footer_row(outcome5_table, values = footer_values, colwidths = c(1,1,1,1))

outcome5_table
```

## Outcome 7: Increase in clients identified who are victims of GBV.
At baseline, healthcare providers (n = 6) self-reported identifying an average of 1.5 clients who were victims of GBV in the past month. At endline, healthcare providers (n = 16) self-reported identifying an average of 3 clients who were victims of GBV in the past month. These self-reported numbers exceeded indicator seven’s target of a 50% increase in clients identified. The number of providers identifying clients experiencing GBV within the past month increased by 167% and the number of clients identified increased by 100%.

Of those healthcare providers who identified clients within the past month who were victims of GBV, there was improvement on all but one appropriate practice from baseline to endline (table 5). The one practice without improvement, offering validating and supportive statements, was already at 100% at baseline and therefore could not improve further.

```{r}
pract19_clean_vars <- names(clean_data)[str_detect(names(clean_data), "practices_clean_19")]

outcome7pre <- clean_data %>%
  filter(status == "All three" & time_point == 1) %>%
  select(all_of(pract19_clean_vars)) %>%
  na.omit() %>%
  t() %>%
  as.data.frame() %>%
  mutate(percent_correct_pre = (rowSums(.))/6*100) %>%
  select(percent_correct_pre)
outcome7pre$question <- rownames(outcome7pre)

outcome7post <- clean_data %>%
  filter(status == "All three") %>%
  filter(time_point == 3) %>%
  select(all_of(pract19_clean_vars)) %>%
  na.omit() %>%
  t() %>%
  as.data.frame() %>%
  mutate(percent_correct_post = (rowSums(.))/16*100) %>%
  select(percent_correct_post)
outcome7post$question <- rownames(outcome7post)

outcome7_pre_post <- merge(outcome7pre, outcome7post, by = "question", all = TRUE) %>%
  mutate(difference = percent_correct_post - percent_correct_pre) %>% 
  mutate(question = factor(question, levels = c("practices_clean_19a", 
                                                        "practices_clean_19b",
                                                        "practices_clean_19c",
                                                        "practices_clean_19d",
                                                        "practices_clean_19e",
                                                        "practices_clean_19f",
                                                        "practices_clean_19g",
                                                        "practices_clean_19h",
                                                        "practices_clean_19i"),
                              labels = c("Assessed the immediate level of danger for the woman",
                    "Discussed the options she may have",
                    "Documented domestic violence history and physical examination findings",
                    "Helped the woman to create a plan to increase her and her children’s safety",
                    "Offered validating and supportive statements",
                    "Provided basic information about domestic violence",
                    "Provided education or resource materials about domestic violence",
                    "Referred the woman to support services in the community",
                    "Talked to the woman about her needs")))

numeric_columns <- sapply(outcome7_pre_post, is.numeric)
outcome7_pre_post[numeric_columns] <- round(outcome7_pre_post[numeric_columns], digits = 2)
  
outcome7_table <- flextable(outcome7_pre_post)

outcome7_table <- set_header_labels(
  outcome7_table,
  question = "Action Taken",
  percent_correct_pre = "Baseline",
  percent_correct_post = "Endline",
  difference = "Percent difference")

outcome7_table <- set_caption(outcome7_table, "**Table 5. Outcome 7: Provider Practices in Identifying and Caring for Victims of GBV**")

outcome7_table
```

## Learning Labs
While the above analysis demonstrates the impact of the entirety of the ‘Responding to GBV Learning Labs’ Curriculum, inclusive of the 5-day intensive training and the 14-month Learning Lab intervention, additional analysis was conducted to understand trends between the end of the 5-day intensive training (the post-intensive assessment) and endline (table 6).
```{r}
clean_scores <- clean_scores %>%
  mutate(timepoint_factored = factor(time_point,
                                     levels = c(2, 3),
                                     labels = c("Post-intensive Training", "Endline")))

learning_labs_table <-
  clean_scores %>%
  filter(status == "All three") %>%  
  filter(time_point != 1) %>%
  tbl_summary(
    include = c("outcome4_score", "outcome5_score", "timepoint_factored"), 
    by = c(timepoint_factored),
    type = c(
      outcome4_score, outcome5_score
    ) ~ "continuous",
    label = list(
      outcome4_score ~ "Outcome 4 Score",
      outcome5_score ~ "Outcome 5 Score"
    ),
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"
    ),
    digits = all_continuous() ~ 2
  ) %>%
  add_p() %>%
  bold_p()

learning_labs_table <- learning_labs_table %>%
  modify_caption("**Table 6. Outcome 4 & 5 Scores from Post-intensive Training to Endline**")

learning_labs_table
```

# Discussion 
Improvements were observed across all MEL indicators, encompassing outcomes 4, 5, and 7, from baseline to endline. These results underscore the positive impact of the 'Responding to GBV' curriculum on enhancing the knowledge and confidence of healthcare providers in delivering care to GBV survivors. This is especially significant in light of the high incidence of GBV in the intervention municipalities and the limited prior GBV training received by participants.

However, it is essential to acknowledge substantial variability in the extent of improvement among individual facilities. Future investigations may benefit from scrutinizing facility characteristics to discern any potential associations between facility or individual provider attributes and varying levels of improvement. Notably, participants from Ermera exhibited lower outcome 4 scores at endline compared to baseline, albeit with a small sample size (two participants). This decline, particularly in domains related to helpful responses, unacceptability of partner violence, and attitudes toward gender roles, raises questions about knowledge retention, potentially influenced by participation in the learning labs, though individual attendance data are unavailable. Further exploration of these subdomains across all participants is warranted to address any issues related to survey questions or translation.

Project records indicate that attendance at monthly learning labs fluctuated between 6-38% of those who attended the intensive training, with 46 providers who attended the intensive training never attending learning labs. The lower attendance rate of the monthly learning labs may explain the plateauing of some participant scores between the post-intensive and endline assessments. Understanding facilitators and barriers to provider attendance at the learning labs could provide valuable insights for future implementation of the 'Responding to GBV' curriculum.

While the number of providers identifying GBV victims in the past month was relatively small at both baseline and endline, significant improvements were observed in this metric and in the use of appropriate practices for these women. The most substantial changes from baseline to endline were noted in referring women to community support services, assisting them in safety planning for themselves and their children, and engaging in discussions about their needs. Notably, six out of nine appropriate practices for women experiencing GBV were reportedly used by 100% of providers at endline. This suggests that advancements in GBV knowledge and confidence translate into tangible improvements in clinical practice, positively affecting women and children affected by GBV. This is particularly noteworthy in an environment where fewer than 9% of training participants had previous GBV training, underscoring the potential impact of the 'Responding to GBV' curriculum on clinical practices.

While this report primarily focuses on changes in provider GBV knowledge and confidence during the project, there is a notable interest in understanding the impact of the learning labs on provider knowledge and confidence. Following the intensive training, participants who attended the learning labs continued to exhibit improvements in outcomes 4 and 5, although outcome 4 did not reach statistical significance. Unfortunately, reliable attendance data are lacking, preventing a comprehensive understanding of how attendance at monthly learning labs may influence score improvements. The sustained score improvements from post-intensive training to endline suggest that the learning labs played a role in maintaining existing knowledge and further enhancing it after the intensive training. However, without a counterfactual, establishing causality is challenging. Investigating the effectiveness of learning labs post-intensive training and assessing the impact of attendance on scores present avenues for future research.

# Limitations:
It is important to acknowledge several limitations in this work:
- Sample Size: Some analyses are constrained by small sample sizes, particularly in Ermera, which may limit the generalizability of the findings.
- Data on Attendance: The absence of individual-level data on attendance at learning labs hinders a thorough assessment of its impact on knowledge and confidence scores.
- Causality: The study's observational nature prevents us from establishing causality between participation in the 'Responding to GBV' curriculum and improvements in knowledge and confidence.
- Survey Questions and Translation: Further investigation is needed to address potential issues related to survey questions or translation, particularly in subdomains showing variation in scores.
- Counterfactual: The lack of a counterfactual group makes it challenging to attribute all observed improvements solely to the curriculum.

Despite these limitations, the study offers valuable insights into the impact of the 'Responding to GBV' curriculum on healthcare providers' knowledge, confidence, and practices in responding to GBV

# Appendix 
## Participant Scores without Facility Categories
```{r}
figure_fp <- paste(gbv_project_wd, "/figures/mean_scores_bar_chart.png", sep = "")
knitr::include_graphics(figure_fp)
```

## Mean Score Differences by Facility
```{r}
figure_fp <- paste(gbv_project_wd, "/figures/mean_diff_bar_plot.png", sep = "")
knitr::include_graphics(figure_fp)
```